---
title: "kaggle_datadiscovery"
author: "Holler Zsuzsa"
date: "January 30, 2016"
output: pdf_document
---

Content summarized in 59 features of all articles published from January 7 2013 to January 7 2015 on Mashable. 39,000 articles in total.

```{r, echo=FALSE, warning=FALSE}
source("/home/zsuzsa/Documents/kaggle/kaggle-onlinenewspopularity/Zsuzsi/data_read.R")
library("ggplot2")
```

##Data discovery

```{r, echo=FALSE}
for (var in c(x.numeric,x.rate)) {
  print(boxplot(data[,var],main=var))
} 

#articles without words
dim(data[data$average_token_length==0,])

#Outlier removed
boxplot(data$n_unique_tokens[data$n_unique_tokens != 701])
boxplot(data$n_non_stop_unique_tokens[data$n_unique_tokens != 701])
boxplot(data$n_non_stop_words[data$n_unique_tokens != 701])

table = table(data.knn[,y],data.knn[,"data_chanel"])
table = round((table * 
                 matrix(1/colSums(table),ncol = ncol(table), nrow=nrow(table) ,byrow=T))
              *100,0)

most.frequent = matrix(NA, 2,7)
frequency = matrix(NA, 2,7)
for (i in 1:2) {
  for (j in 1:7) {
    most.frequent[i,j] = 
      which.max(table(data[,y],data[,"data_chanel"],data[,"is_weekend"])[,j,i])
    frequency[i,j] = 
      max(table(data[,y],data[,"data_chanel"],data[,"is_weekend"])[,j,i])
  }
}


qplot( factor(popularity), data=data, geom = "bar",
       fill=factor(weekday)
       #position = "fill" #different display
)

ggplot( data , aes(x = factor(popularity) )) +  
        geom_bar(aes(y = (..count..)/sum(..count..),fill = factor(weekday)), 
                 binwidth = 25) + 
        scale_y_continuous(labels = scales::percent) +
        geom_text(aes(y = ((..count..)/sum(..count..)),
                      label=scales::percent((..count..)/sum(..count..))),
                  stat="bin",vjust = -0.25)
      
qplot( factor(popularity), data=data, geom="freqpoly", 
       group=factor(weekday), 
       colour=factor(weekday),
       position="identity"
)

ggplot( data , aes(x = factor(popularity) )) +  
        facet_grid(.~data_chanel) +
        geom_histogram()

```

###Explore relationship between numeric data
Simple correlations for non-categorical variables visualised. Stronger correlations between variables of the same category.

```{r, echo=FALSE}
#Correlations for numeric data
library(reshape2)

corr <-round(cor( data[,sort(c(x.numeric,x.rate))], method="pearson", use="pairwise.complete.obs"),2)
corr[lower.tri(corr)] <- NA
corr <- melt(corr)
corr$Var1<-as.character(corr$Var1)
corr$Var2<-as.character(corr$Var2)
corr <- na.omit(corr)

ggplot(corr, aes(Var2, Var1))+
  geom_tile(data=corr, aes(fill=value), color="white")+
  scale_fill_gradient2(low="blue", high="red", mid="white", 
                       midpoint=0, limit=c(-1,1),name="Correlation\n(Pearson)")+
  theme(axis.text.x = element_text(angle=45, vjust=1, size=11, hjust=1))+
  coord_equal()

```

###Explore relationships between categorical variables (including our y variable.)

```{r, echo=FALSE, warning=FALSE}

library(MASS)
library(entropy)

#contingency tables
tbl.channel = table( data[,y] , data[,"data_chanel"]) 
freq.tbl.channel = tbl.channel*100/sum(tbl.channel)

tbl.weekday = table( data[,y] , data[,"weekday"]) 
freq.tbl.weekday = tbl.weekday*100/sum(tbl.weekday)
```

Mosaic plot: Each cell is represented by a box whose area is equal to the count in that cell. The width of the box is the same for all boxes in the same column and is equal to the total count in that column. The height of the box is the proportion of individuals in the column which fall into that cell. (Empirical conditional probabilities of row variable given column variable.)

Colored by: Pearson residuals - departure of each cell from independence. The formula is (actual - expected)/sqrt(expected). The units are in standard deviations, so a residual greater than 2 or less than -2 represents a departure significant at the 95% level. The expected count under independence is (row marginal)*(column marginal)/(table total).

```{r, echo=FALSE, warning=FALSE}
mosaicplot(t(tbl.channel), shade=T)

```

```{r, echo=FALSE, warning=FALSE}
mosaicplot(t(tbl.weekday), shade=T)
```

Chi-squared test/Fisher test/mutual information for categorical data:

Fisher test - were not able to perform - requires a lot of memory - very similar to chi-squared

```{r, echo=FALSE, warning=FALSE}
by.cat <- function(cont.table){
  by.cat = data.frame(chisq = c(),mi = c(), mi.sd = c() ,fisher=c())
  n.cat = ncol(cont.table)
  for (i in 1:n.cat) {
    ctbl = cbind(cont.table[,i], rowSums(cont.table[,-i])) 
    by.cat[i,"chisq"] = chisq.test(ctbl)$p.value
    by.cat[i,"mi"] = mi.empirical(ctbl)
    by.cat[i,"mi.sd"] = mi.empirical(ctbl)/entropy.empirical( colSums(ctbl) )
    rownames(by.cat)[i]=colnames(cont.table)[i]
  }
  return(by.cat)
}
```

Channel categories in the order of comovement by chi-squared, mutual information and standardised mutual information.

```{r, echo=FALSE, warning=FALSE}
channel.by.cat = by.cat(tbl.channel)
rownames(channel.by.cat[order(channel.by.cat$chisq),])
rownames(channel.by.cat[order(channel.by.cat$mi,decreasing = T),])
rownames(channel.by.cat[order(channel.by.cat$mi.sd, decreasing = T),])
```

Days of week in the order of comovement by chi-squared, mutual information and standardised mutual information.

```{r, echo=FALSE, warning=FALSE}
weekday.by.cat = by.cat(tbl.weekday)
rownames(weekday.by.cat[order(weekday.by.cat$chisq),])
rownames(weekday.by.cat[order(weekday.by.cat$mi,decreasing = T),])
rownames(weekday.by.cat[order(weekday.by.cat$mi.sd,decreasing = T),])

```

```{r, echo=FALSE, warning=FALSE}
#Digital variables
vp.setup(1,2)
for (i in 1:length(x.dig)) {
  print( ggplot( data , aes_string(x=x.dig[i]) ) + geom_histogram(binwidth=1 ), vp=vp.layout(1,i))
}

vp.setup(1,2)
for (i in 1:length(x.dig)) {
  print(ggplot(data = data, aes_string(x=x.dig[i], y="popularity") ) +
  geom_smooth(se=FALSE) +
  geom_point( alpha=0.5), vp=vp.layout(1,i))
}

vp.setup(1,2)
for (i in 1:length(x.dig)) {
  print(ggplot(data = data, aes_string(x=x.dig[i], y="popularity") ) +
  geom_smooth(aes(color=as.factor(data_chanel)),se=FALSE) +
  geom_point( aes(color=as.factor(data_chanel)),alpha=0.5), vp=vp.layout(1,i))
}

vp.setup(1,2)
for (i in 1:length(x.dig)) {
  print(ggplot(data = data, aes_string(x=x.dig[i], y="popularity") ) +
  geom_smooth(aes(color=as.factor(is_weekend)),se=FALSE) +
  geom_point( aes(color=as.factor(is_weekend)),alpha=0.5), vp=vp.layout(1,i))
}

#Links
vp.setup(2,3)
for (i in 1:length(x.links)) {
  row = ceiling(i/3)
  col = i - (row-1)*3
  print(ggplot( data , aes_string(x=x.links[i]) ) + geom_histogram(),vp=vp.layout(row,col))
}

vp.setup(2,3)
for (i in 1:length(x.links)) {
  row = ceiling(i/3)
  col = i - (row-1)*3
  print(ggplot(data = data, aes_string(x=x.links[i], y="popularity") ) +
  geom_smooth(se=FALSE) +
  geom_point( alpha=0.5), vp=vp.layout(row,col))
}

#Words
vp.setup(2,3)
for (i in 1:length(x.words)) {
  row = ceiling(i/3)
  col = i - (row-1)*3
  print(ggplot( data , aes_string(x=x.words[i]) ) + geom_histogram(),vp=vp.layout(row,col))
}

#Key
vp.setup(2,5)
x.temp = x.key[!(x.key %in% x.cat)]
for (i in 1:length(x.temp)) {
  row = ceiling(i/5)
  col = i - (row-1)*5
  print(ggplot( data , aes_string(x=x.temp[i]) ) + geom_histogram(),vp=vp.layout(row,col))
}

####
vp.setup(4,4)
for (var in c(x.numeric ,x.rate)) {
  
}

vp.setup(2,5)
for (i in 1:length(x.temp)) {
  row = ceiling(i/5)
  col = i - (row-1)*5
  print(ggplot(data = data, aes_string(x=var, y="popularity") ) +
  geom_smooth(aes(color=as.factor(data_chanel)),method = "lm") +
  geom_point( aes(color=as.factor(data_chanel)),alpha=0.5) )
}



```
